---
title: "Cyclistic tripdata analysis 2"
author: "Ha Luong"
date: "3/26/2023"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
setwd("~/Documents/Google_Data_Analytics/Course_8/Cyclistic_Trip_Data /Cyclistic_CSV_files")
df <- read_csv("cyclistic_clean.csv")
```

# Analyze the difference in bike usage between casual and annual members
## Some statistics
```{r}
options(scipen = 999)
ggplot(df, aes(x=member_casual, fill=member_casual)) +
    geom_bar() +
    geom_text(df,mapping = aes(label=..count..),stat = "count",position = "stack", vjust=-0.3) +
    labs(x="Customer types", y="Total number of customers", title="Casuals and Members distribution") 
```

## Start hour throughout the week
```{r}
ggplot(data =df) +
  geom_bar(mapping=aes(x = starting_hour, fill = member_casual)) +
  facet_wrap(~day_of_week) +
  labs(x = 'Starting hour', y = 'Number of rides', fill = 'Member type', title = 'Hourly use of bikes throughout the week') +
  theme(axis.text = element_text(size = 5))
ggsave("Hourly_use_of_bikes_throughout_the_week.png", dpi = 1000)
```
## Analyze daily usage 
### We need to assign chronological order for the day of week so that the plot will show in chronological order. I use function "options(scipen = 999)" to extends the number of lines of printing so that the number of ride can be displayed in normal number intead of scientific number

```{r}
df$day_of_week <- factor(df$day_of_week, levels = c("Sun", "Mon","Tue", "Wed", "Thu", "Fri", "Sat"))
options(scipen = 999)
ggplot(data = df) +
  aes(x = day_of_week, fill = member_casual) +
  geom_bar(position = 'dodge') +
  labs(x = 'Day of week', y = 'Number of rides', fill = 'Member type', title = 'Number of rides by member type during a week')
ggsave("daily_ride_by_member_type.png")
```
## Analyze by monthly usagge
```{r}
options(scipen = 999)
ggplot(data = df) +
  aes(x = month, fill = member_casual) +
  geom_bar(position = 'dodge') +
  labs(x = 'Month', y = 'Number of rides', fill = 'Member type', title = 'Monthly rides by member type')
ggsave("monthly_rides_by_member_type.png")
```
## Average trip duration by day of the week
```{r}
ggplot(df, aes(x = factor(day_of_week), y=ride_length)) + 
  geom_bar(stat = "summary", fun = "mean", fill="#56B4E9") +
  facet_wrap(~member_casual)+
  labs(x = 'Day of week', y = 'Average ride length (min)', fill = 'Member type', title = 'Average ride length by day of the week')
ggsave("Ride_length.png")
```
## Average trip duration by month
```{r}
ggplot(df, aes(x = factor(month), y=ride_length)) + 
  geom_bar(stat = "summary", fun = "mean", fill="#56B4E9") +
  facet_wrap(~member_casual)+
  labs(x = 'Month', y = 'Average ride length (min)', fill = 'Member type', title = 'Average ride length by month')
ggsave("Ride_length_by_month.png")
```
## Average temparature of Chicago (Chicago average daily mean tempreture (Midway Airport), 1991–2020 normals is from: https://en.wikipedia.org/wiki/Climate_of_Chicago)
```{r}
chicago_mean_temp <- c(-3.2, -1.2, 4.4, 10.5, 16.6, 22.2, 24.8, 23.9, 19.9, 12.9, 5.8, -0.3)
month <- c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec")
chicago_temp_df <- data.frame(chicago_mean_temp, month)
chicago_temp_df$month <- factor(chicago_temp_df$month, levels = c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec"))
ggplot(chicago_temp_df, aes(x=month, y=chicago_mean_temp)) +
    labs(x="Month", y="Mean temperature", title="Daily mean temperature for Chicago (1991-2020) in Celsius") +
    geom_col(, fill="#56B4E9")
ggsave("chicago_daily_temp.png")
```
## Analyze by start station and end station
### Number of rides by start station
```{r}
library(dplyr)
library(data.table)
df1 <- df %>% group_by(start_station_name, member_casual) %>% tally(sort = TRUE)
df1<- na.omit(df1)
setDT(df1)
df1[, as.list(summary(n)), by = member_casual]
```
### Number of rides by end station
```{r}
df2 <- df %>% group_by(end_station_name, member_casual) %>% tally(sort = TRUE)
df2<- na.omit(df2)
setDT(df2)
df2[, as.list(summary(n)), by = member_casual]
```
