---
title: "Google Data Anaylystics Capstone Project"
author: "Ha Luong"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
#install.packages("tidyverse")
library(tidyverse)
#install.packages("skimr")
library(skimr)
#install.packages("janitor")
library(janitor)
knitr::opts_chunk$set(echo = TRUE)
```

# Open file and assign name to each files
```{r}
setwd("~/Documents/Google_Data_Analytics/Course_8/Cyclistic_Trip_Data /Cyclistic_CSV_files")
Mar_2022 <- read_csv("2022_03.csv")
Apr_2022 <- read_csv("2022_04.csv")
May_2022 <- read_csv("2022_05.csv")
Jun_2022 <- read_csv("2022_06.csv")
Jul_2022 <- read_csv("2022_07.csv")
Aug_2022 <- read_csv("2022_08.csv")
Sep_2022 <- read_csv("2022_09.csv")
Oct_2022 <- read_csv("2022_10.csv")
Nov_2022 <- read_csv("2022_11.csv")
Dec_2022 <- read_csv("2022_12.csv")
Jan_2023 <- read_csv("2023_01.csv")
Feb_2023 <- read_csv("2023_02.csv")
```

# Check for column name and format type
```{r Check for column name and format type}
str(Mar_2022)
str(Apr_2022)
str(May_2022)
str(Jun_2022)
str(Jul_2022)
str(Aug_2022)
str(Sep_2022)
str(Oct_2022)
str(Nov_2022)
str(Dec_2022)
str(Jan_2023)
str(Feb_2023)
```

Combine 12 tables into 1 table called merged_df

```{r Merge all datasets into one table}
merged_df <- bind_rows(Mar_2022, Apr_2022, May_2022, Jun_2022, Jul_2022, Aug_2022, Sep_2022, Oct_2022, Nov_2022, Dec_2022, Jan_2023, Feb_2023)
```

# Create new column to analyze by different categories
### Start hour
```{r}
merged_df$starting_hour <- format(as.POSIXct(merged_df$started_at), '%H')
```
### Day of the week
```{r}
merged_df$day_of_week <- wday(merged_df$started_at, label = T, abbr = T)
```
### Month
```{r}
merged_df$month <- format(as.Date(merged_df$started_at), '%m')
```
### Trip Duration
```{r}
merged_df$ride_length <- difftime(merged_df$ended_at, merged_df$started_at, units ='min')
merged_df$ride_length <- as.numeric(merged_df$ride_length)
```

# Cleanning data
### Check for format of new column
```{r}
str(merged_df)
```
### Summary of the data to see any inappropriate data
```{r}
summary(merged_df)
```

### Cleaning & removing any spaces, parentheses, etc in the column names
```{r Cleaning & removing any spaces, parentheses, etc}
merged_df <- clean_names(merged_df)
```
### Removing Duplicates
```{r}
merged_df %>%  distinct()
```
### Check for mean and sd of ride_length
```{r}
sd(merged_df$ride_length)
mean(merged_df$ride_length)
```
### Remove any rows which have ride_length <= 0 and ride_length > 544.6 mins
```{r}
cleaned_df <- merged_df[!(merged_df$ride_length>544.6) & !(merged_df$ride_length<=0),]
```

### Summary of the cleaned data
```{r}
summary(cleaned_df)
```
# Save as new file for analysis

```{r}
cleaned_df %>%
  write.csv("cyclistic_clean.csv")
```